\documentclass[twosided,openany]{book}
\input{surimpreamble}
\begin{document}
\title{SURIM 2012 Lecture Notes}
\author{\href{mailto:adebray@stanford.edu}{Arun Debray}}
\maketitle
\tableofcontents
\chapter{Explorations in Number Theory}
    \section{Class Numbers of Quadratic Forms: June 25, 2012}
	\input{classnumbers}
    \section{Quadratic Reciprocity: June 26, 2012}
	\input{quadraticreciprocity}
    \section{Linear Recurrences: June 26, 2012}
	\input{linearrecurrences}
    \section{Linear Recurrences, Part II: June 28, 2012}
	\input{linear2}
    \section{The Skolem-Mahler-Lech Theorem: June 29, 2012}
	\input{skolem}
    \section{People Care About Class Numbers: July 3, 2012}
    	\input{useful}
    \section{Quadratic Fields: July 5, 2012}
	\input{quadraticfields}
    \section{Some Group Theory Preliminaries: July 17, 2012}
	\input{grouptheory}
    \section{Some Background on Groups: July 24, 2012}
	\input{grouptheory2}
    \section{Computations in Quadratic Fields: July 26, 2012}
	\input{idealstuff}
    \section{Short Exact Sequences: July 31, 2012}
	\input{exact}
    \section{Finitely Generated Abelian Groups: July 31, 2012}
	\input{abeliandecomp}
    \section{The Cohen-Lenstra Heuristics: August 10, 2012}
	\input{cohen}
    \section{The Analytic Class Number Formula: August 16, 2012}
	\input{formula}
\chapter{Mathematical Lectures}
    \section{Factoring, by Dr. Akshay Venkatesh: June 27, 2012}
	\input{factoring}
    \section{{\sc Sage} in Brief, by Dr. Justin Walker: July 9, 2012}
	\input{sage}
    \section{The Geometry of Numbers, by Dr. Brian Conrad: July 11, 2012}
	\input{minkowski}
    \section{Game Theory, by Michael Manapat: July 18, 2012}
	\input{gametheory}
    \section{Dispersal and Aggregation, by Dr. Nancy Rodriguez: July 25, 2012}
	\input{pde}
    \section{Earthquake Cycle Simulations, by Dr. Brittany Erickson: July 30, 2012}
	\input{geology}
    \section{Some Differential Geometry, by Dr. Rafe Mazzeo: August 1, 2012}
	\input{diffgeo}
    \section{Hopf Algebras and {\sc Sage}, by Dr. Daniel Bump: August 8, 2012}
	\input{hopf}
    \section{The Shape of Data, by Dr. Gunnar Carlsson: August 13, 2012}
	\input{shapeofdata}
    \section{Mathematical Doodling, by Dr. Ravi Vakil: August 15, 2012}
	\input{doodle}
\chapter{Lectures About Research}
    \section{Effective Mentor-Mentee Relationships, by Brian Thomas: June 29, 2012}
	\input{mentor}
    \section{How to Give an Effective Presentation, by Joyce Moser: July 5, 2012}
	\input{presentation}
	%Time to study some presentation theory
\chapter{Lectures by the Statistics Department}
    \section{RStudio and \LaTeX{}: June 28, 2012}
	\input{unix}
    \section{A Short Introduction to Unix: June 28, 2012}
	\input{unix2}
    \section{Plotting in R using \texttt{ggplot2}: July 3, 2012}
	\input{morer}
    \section{File I/O and Functions: July 5, 2012}
	\input{fileio}
    \section{A Brief Introduction to \LaTeX{}: July 10, 2012}
	\input{la}
    \section{Control Sequences in R: July 12, 2012}
	\input{rcontrol}
    \section{How to Make Presentations: July 17, 2012}
	\input{statspresent}
    \section{The Singular Value Decomposition: July 19, 2012}
	\label{svd}
	\input{svd}
    \section{Correspondence Analysis: July 24, 2012}
	\input{datamining}
\end{document}
